Code for *Tree-Ring Evidence of Increasing Drought Risks amidst Projected Flood Intensification in the Kabul River Basin (Afghanistan and Pakistan)* by Khan et al. (2022)

### 2022-10-21 paper-code.Rmd
Revised after peer review.

```{r, message=FALSE, warning=FALSE}
library(dplR)       # Tree ring data processing
library(ldsr)       # Tree ring data processing
library(data.table) # Data handling 
library(missMDA)    # Imputation
library(qmap)       # Bias correction
library(modifiedmk) # Trend analysis
library(ggplot2)    # Plotting
library(cowplot)    # Plotting
library(patchwork)  # Plotting
library(ggprism)    # Plotting
library(ggnewscale) # Plotting
theme_set(theme_prism(base_size = 10, base_fontface = 'plain', base_line_size = 0.2))
```

@@

```{r, eval=FALSE}
rmarkdown::render('paper-code.Rmd', output_format = 'all', output_options = list(hightlight = 'tango'))
```

@@

@@

@@

**Figure S1 - Chronology and SSS plot**

@@

**Figure S2**

@@

```{r}
PCfit <- prcomp(X, scale. = TRUE)
PC <- PCfit$x
pcDT <- as.data.table(as.data.frame(summary(PCfit)$importance),
                      keep.rownames = 'var') |> 
  melt(id.vars = 'var', variable.name = 'PC')
```

```{r, fig.width=7, fig.height=4}
p1 <- ggplot(pcDT[var == 'Standard deviation']) +
  geom_col(aes(PC, value), color = 'black', fill = 'gray95') +
  geom_hline(yintercept = 1, size = 0.8) +
  scale_y_continuous(expand = expansion(add = c(0, 0.01)),
                     limits = c(0, 2)) +
  labs(x = NULL, y = 'Eigenvalue [-]', tag = 'a)') +
  theme(
    panel.grid.major.y = element_line('gray'),
    panel.ontop = TRUE,
    axis.ticks = element_blank(),
    axis.line = element_blank())
p2 <- ggplot() +
  geom_col(
    aes(PC, value * 100), 
    pcDT[var == 'Proportion of Variance'],
    fill = 'steelblue', color = 'black', alpha = 0.8) +
  geom_line(
    aes(PC, value * 100, group = 1),
    pcDT[var == 'Cumulative Proportion'],
    color = 'firebrick', size = 0.8) +
  geom_point(
    aes(PC, value * 100),
    pcDT[var == 'Cumulative Proportion'],
    color = 'firebrick', size = 1.6) +
  labs(x = NULL, y = 'Proportion of variance [%]', tag = 'b)') +
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  theme(
    panel.grid.major.y = element_line('gray'),
    panel.ontop = TRUE,
    axis.ticks = element_blank(),
    axis.line = element_blank())

p1 + p2
```

@@

@@

```{r}
round(lmCV$metrics[, 1:3], 2)
```
**Figure S4 - score distribution**

@@

Calculate rolling statistics and extract drought events.

```{r}
alignType <- 'right'
recFinal[, rolMax := frollapply(Qbc, 50, max, align = alignType)]
recFinal[, rolMin := frollapply(Qbc, 50, min, align = alignType)]
recFinal[, rolMed := frollapply(Qbc, 50, median, align = alignType)]
recFinal[, period := fcase(
  year %in% 1517:1767, 1,
  default = 2)]
densCals <- recFinal[, {
  d <- density(Qbc, cut = 0, bw = 45)
  list(x = d$x, y = d$y / max(d$y) * 1.5)
}, by = period]

medians <- recFinal[, {
  d <- density(Qbc, cut = 0, bw = 45)
  m <- median(Qbc)
  y <- approx(d$x, d$y, m)$y
  list(x = m, y = y / max(d$y) * 1.5)
}, by = period]

@@

Mann-Kendall trend test with trend-free pre-whitening.

```{r}
recFinal[!is.na(rolMax), tfpwmk(rolMax)] |> round(6)
recFinal[!is.na(rolMin), tfpwmk(rolMin)] |> round(6)
recFinal[!is.na(rolMed), tfpwmk(rolMed)] |> round(6)
```

@@

```{r, fig.width=8, fig.height=6, message=FALSE}
p1 <- ggplot(recFinal) +
  labs(x = NULL, y = 'P [mm]') +
  geom_line(aes(year, rolMax, colour = '50-yr max'), na.rm = TRUE, size = 0.4, alpha = 0.5) +
  geom_line(aes(year, rolMin, colour = '50-yr min'), na.rm = TRUE, size = 0.4, alpha = 0.5) +
  geom_line(aes(year, rolMed, colour = '50-yr median'), na.rm = TRUE, size = 0.4, alpha = 1) +
  geom_line(aes(year, Qbc, color = 'Annual')) +
  geom_smooth(aes(year, rolMax, colour = '50-yr max'), size = 0.6,
              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
  geom_smooth(aes(year, rolMin, colour = '50-yr min'), size = 0.6,
              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
  geom_smooth(aes(year, rolMed, colour = '50-yr median'), size= 0.6,
              formula = 'y ~ x', method = 'lm', na.rm = TRUE, fill = NA) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    minor_breaks = seq(200, 900, 50),
    breaks = seq(200, 900, 100),
    limits = c(200, 900),
    labels = skip_label(2),
    guide = guide_prism_offset_minor()) +
  labs(x = NULL, y = 'Precipitation [mm]', colour = NULL, tag = 'a)') +
  scale_color_manual(values = c(RColorBrewer::brewer.pal(3, 'Set2'), 'gray')) +
  theme(
    legend.key.width = unit(0.5, 'cm'),
    legend.position = 'top')

p2 <- ggplot(densCals) +
  geom_ribbon(aes(x, ymin = period, ymax = y + period, group = factor(period),
                  fill = factor(period)), 
              alpha = 0.5) +
  geom_line(aes(x, y + period, group = factor(period))) +
  geom_linerange(aes(x, ymin = period, ymax = period + y), medians, colour = 'firebrick') +
  scale_fill_manual(
    labels = c('1517-1717', '1718-2018'),
    values = wesanderson::wes_palette('Royal1', 4)[c(1, 4)]) +
  scale_x_continuous(
    guide = guide_prism_offset_minor(),
    breaks = seq(200, 900, 100), 
    minor_breaks = seq(200, 900, 20)) +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = c(1.5, 2.5),
    labels = c('1517-1717', '            1718-2018')) +
  labs(x = 'Precipitation [mm]', y = NULL, tag = 'b)') +
  theme(
    legend.position = 'none',
    axis.line.y = element_blank(),
    axis.text.y.left = element_text(angle = 90, hjust = 0.5),
    axis.ticks.y = element_blank())

@@

>>>>

```{r, message=FALSE, warning=FALSE}
library(dplR)         # Tree ring data processing
library(ldsr)         # Tree ring data processing
library(data.table)   # Data handling 
library(missMDA)      # Imputation
library(qmap)         # Bias correction
library(quantreg)     # Quantile regression
library(ggplot2)      # Plotting
library(cowplot)      # Plotting
library(patchwork)    # Plotting
library(ggprism)      # Plotting
library(ggnewscale)   # Plotting
library(RColorBrewer) # Color palette
theme_set(theme_prism(base_size = 10, base_fontface = 'plain', base_line_size = 0.2))
```

@@

```{r, eval=FALSE}
rmarkdown::render(
  'paper-code.Rmd', 
  output_format = 'all', 
  output_options = list(hightlight = 'tango', 
                        latex_engine = 'xelatex'))
```

@@

### Precipitation and Temperature

@@

### Wind vectors

Note: reanalysis and GPCP data are huge so they won't be included in this repository. You can download them from NOAA in order to run the code below, which produces Figure 1d.

```{r, eval=FALSE}
# Set map limit
xrange <- c(0, 80)
yrange <- c(0, 50)

world <- map('world', fill = TRUE, plot = FALSE) |> 
  st_as_sf(coords = c('x', 'y')) |> 
  st_buffer(0.01) |>
  st_union()

pak <- map('world', 'Pakistan', fill = TRUE, plot = FALSE) |> 
  st_as_sf(coords = c('xy', 'y'))

## NCEP wind data
# Wind vector: read u-wind and v-wind, then merge
# Change the file path to your own
ncep <- merge(
  # u-wind
  ReadNetCDF('data/uwnd.mon.mean.nc', 'uwnd', 
             subset = list(level = 850, 
                           lon = xrange, lat = yrange,
                           time = c('1979-01-01', '2021-12-01'))),
  ReadNetCDF('data/vwnd.mon.mean.nc', 'vwnd', 
             subset = list(level = 850, 
                           lon = xrange, lat = yrange,
                           time = c('1979-01-01', '2021-12-01'))),
  by = c('time', 'level', 'lat', 'lon'))
# Create seasonal mean. 
ncepSeas <- ncep[month(time) %in% c(2:4, 6:8), 
                 .(uwnd = mean(uwnd), vwnd = mean(vwnd)), 
                 by = .(lon, lat, seas = fifelse(month(time) %in% 2:4, 'Feb-Apr', 'Jun-Aug'))]

## GPCP precipitation data
pre <- ReadNetCDF('data/precip.mon.mean.nc', 'precip',
                  subset = list(lon = xrange, lat = yrange,
                           time = c('1979-01-01', '2015-12-01')))
# Create seasonal mean. 
preSeas <- pre[month(time) %in% c(2:4, 6:8), 
               .(precip = mean(precip)), 
               by = .(lon, lat, seas = fifelse(month(time) %in% 2:4, 'Feb-Apr', 'Jun-Aug'))]
```

```{r wind plot, fig.width=7, fig.height=3.5, eval=FALSE}
maxSpeed <- ncepSeas[, sqrt(uwnd^2 + vwnd^2) |> max() |> ceiling()]
maxRain <- preSeas[, ceiling(max(precip))]
p3 <- ggplot() +
  # Precip
  geom_contour_fill(
    aes(lon, lat, z = precip, fill = stat(level_d)),
    data = preSeas,
    breaks = seq(0, maxRain, 2)) +
  # Land mass
  geom_sf(data = world, size = 0.2, fill = NA) +
  geom_sf(data = pak, size = 0.4, fill = NA, color = 'firebrick') +
  # Arrow vector
  geom_arrow(aes(lon, lat, dx = uwnd, dy = vwnd), 
             data = ncepSeas,
             # min.mag = 1,
             # skip = 1,
             color = blues9[7],
             size = 0.2,
             arrow.length = 0.3) +
  # Site locations
  geom_point(aes(Longitude, Latitude), crnMeta, color = 'darkviolet') +
  # Facet by seasons
  facet_wrap(vars(seas)) +
  # Set scales
  as.discretised_scale(scale_fill_distiller)(
    name = 'Precipitation rate [mm/day]',
    palette = 'Greens', direction = 1, breaks = seq(0, maxRain, 2),
    guide = guide_colorsteps(
      # override.aes = list(alpha = 0.5),
      title.position = 'top', 
      show.limits = TRUE)) + 
  scale_x_continuous(expand = c(0, 0), labels = paste_long) +
  scale_y_continuous(expand = c(0, 0), labels = paste_lat) +
  scale_mag(
    name = 'Wind speed [m/s]', 
    guide = guide_vector(title.position = 'top'),
    max = 20) +
  # guides(alpha = guide_none()) +
  # Axis labels
  labs(x = NULL, y = NULL, tag = 'd)') +
  # Map projection
  coord_sf(xlim = xrange + c(1, -1.25), ylim = yrange + c(1.25, 0)) +
  panel_border('black', 0.2) +
  theme(
    strip.text = element_text(face = 'bold'),
    legend.position = 'bottom',
    legend.key.width = unit(1.25, 'cm'),
    legend.title = element_text()) 
p3
# ggsave('precip-wind-pakistan.png', width = 7, height = 3.5)
```

@@

**Figure S2 - Chronology and SSS plot**

@@

**Figure S5**

@@

```{r}
PCfit <- prcomp(X, scale. = TRUE)
PC <- PCfit$x
pcDT <- as.data.table(as.data.frame(summary(PCfit)$importance),
                      keep.rownames = 'var') |> 
  melt(id.vars = 'var', variable.name = 'PC')

# Extract PC1 and PC2 for plotting
pc12plot <- as.data.table(PC[, 1:2, drop = FALSE])[, year := crnWide$year]
```

**Figure S6**

```{r, fig.width=7, fig.height=6}
p1 <- ggplot(pcDT[var == 'Standard deviation']) +
  geom_col(aes(PC, value), color = 'black', fill = 'gray95') +
  geom_hline(yintercept = 1, size = 0.8) +
  scale_y_continuous(expand = expansion(add = c(0, 0.01)),
                     limits = c(0, 2)) +
  labs(x = NULL, y = 'Eigenvalue [-]', tag = 'a)') +
  theme(
    panel.grid.major.y = element_line('gray'),
    panel.ontop = TRUE,
    axis.ticks = element_blank(),
    axis.line = element_blank())
p2 <- ggplot() +
  geom_col(
    aes(PC, value * 100), 
    pcDT[var == 'Proportion of Variance'],
    fill = 'steelblue', color = 'black', alpha = 0.8) +
  geom_line(
    aes(PC, value * 100, group = 1),
    pcDT[var == 'Cumulative Proportion'],
    color = 'firebrick', size = 0.8) +
  geom_point(
    aes(PC, value * 100),
    pcDT[var == 'Cumulative Proportion'],
    color = 'firebrick', size = 1.6) +
  labs(x = NULL, y = 'Proportion of variance [%]', tag = 'b)') +
  scale_y_continuous(expand = expansion(add = c(0, 1))) +
  theme(
    panel.grid.major.y = element_line('gray'),
    panel.ontop = TRUE,
    axis.ticks = element_blank(),
    axis.line = element_blank())

p3 <- ggplot(pc12plot) +
  geom_line(aes(year, PC1), color = blues9[9]) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_minor()) +
  labs(x = NULL, tag = 'c)')

p4 <- ggplot(pc12plot) +
  geom_line(aes(year, PC2), color = blues9[9]) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_minor()) +
  labs(x = NULL, tag = 'd)')

# layout <- '
# AB
# CC
# DD
# '

layout <- c(
  patchwork::area(1, 1, 3, 1),
  patchwork::area(1, 2, 3, 2),
  patchwork::area(4, 1, 5, 2),
  patchwork:: area(6, 1, 7, 2)
)
p1 + p2 + p3 + p4 +
  plot_layout(design = layout)
```

@@

**Figure S1 - Autocorrelation**

```{r}
acf(PSepAug$Qa, main = NA, xlab = 'Lag [years]')
```

@@

```{r}
round(lmCV$metrics[, 1:3], 2)
```
**Figure S7 - score distribution**

@@

## Quantile regression

```{r}
taus <- c(0.01, 0.05, 0.50, 0.95, 0.99)
# names(taus) <- sprintf(taus, fmt = '%.2f')
names(taus) <- paste0(taus * 100, '%')
qrDT <- taus |> 
  lapply(\(tau) {
    fit <- rq(Qbc ~ year, tau, recFinal)
    summ <- summary(fit, se = 'boot')
    data.table(
      b = fit$coefficients[1], 
      a = fit$coefficients[2], 
      signif = summ$coefficients[2, 4] < 0.1)
  }) |> 
  rbindlist(idcol = 'tau')
```

```{r}
recFinal[, period := fcase(
  year %in% 1517:1767, 1,
  default = 2)]
densCals <- recFinal[, {
  d <- density(Qbc, cut = 0, bw = 45)
  list(x = d$x, y = d$y / max(d$y) * 1.5)
}, by = period]

@@

@@

```{r, fig.width=8, fig.height=6, message=FALSE}
reds <- brewer.pal(9, 'Reds')
p1 <- ggplot() +
  geom_line(aes(year, Qbc), recFinal, color = 'gray85') +
  geom_abline(aes(slope = a, intercept = b, color = tau, linetype = signif), 
              qrDT, show.legend = FALSE) +
  geom_line(aes(x = 1800, y = 400, color = tau, group = 1), qrDT) +
  geom_point(aes(year, Qbc), recFinal, color = 'gray50', shape = 1, size = 0.3) +
  scale_color_manual(
    values = c(reds[6], reds[3], 'darkviolet', blues9[4], blues9[7])) +
  scale_linetype_manual(values = c(2, 1)) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    breaks = seq(200, 800, 100),
    minor_breaks = seq(200, 850, 50),
    limits = c(200, 850),
    guide = guide_prism_offset_minor()) +
  guides(color = guide_legend(
    title.position = 'top',
    title.hjust = 0.5,
    override.aes = list(size = 0.6))) +
  labs(x = NULL, y = 'Precipitation [mm]', tag = 'a)') +
  theme(
    plot.tag.position = c(0, 0.94),
    legend.position = 'top')

p2 <- ggplot(densCals) +
  geom_ribbon(aes(x, ymin = period, ymax = y + period, group = factor(period),
                  fill = factor(period)), 
              alpha = 0.5) +
  geom_line(aes(x, y + period, group = factor(period))) +
  scale_fill_manual(
    labels = c('1517-1717', '1718-2018'),
    values = wesanderson::wes_palette('Royal1', 4)[c(1, 4)]) +
  scale_x_continuous(
    guide = guide_prism_offset_minor(),
    breaks = seq(200, 900, 100), 
    minor_breaks = seq(200, 900, 20)) +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = c(1.5, 2.5),
    labels = c('1517-1717', '            1718-2018')) +
  labs(x = 'Precipitation [mm]', y = NULL, tag = 'b)') +
  theme(
    legend.position = 'none',
    axis.line.y = element_blank(),
    axis.text.y.left = element_text(angle = 90, hjust = 0.5),
    axis.ticks.y = element_blank())

@@

# Sensitivity analyses

We now rerun the reconstruction procedure with different SSS cut-off values.

## SSS = 0.75

```{r}
firstYear2 <- crn[sss > 0.75][, .N, by = year][N >= 4, first(year)]
lastYear <- 2018
crnWide2 <- crn[year %in% firstYear2:lastYear,
                dcast(.SD, year ~ site, value.var = c('res'))]
X2 <- crnWide2[, -'year'] |> 
  as.matrix() |> 
  imputePCA(ncp = 6) |> 
  {\(x) x$completeObs}() 
PC2 <- prcomp(X2, scale. = TRUE)$x
sv <- c(1, 2)
sPC2 <- as.data.table(PC2[, sv, drop = FALSE])
lmFit2 <- PCR_reconstruction(PSepAug, sPC2, firstYear2, transform = 'none')
set.seed(42)
lmCV2 <- cvPCR(PSepAug, sPC2, firstYear2, transform = 'none',
              Z = cvFolds, use.robust.mean = FALSE)
recFinal2 <- lmFit2$rec
bcFit2 <- fitQmap(
  PSepAug$Qa,
  recFinal2[year %in% PSepAug$year, Q],
  'RQUANT',
  wet.day = FALSE)
recFinal2[, Qbc := doQmap(Q, bcFit)]
```

## SSS = 0.85

```{r}
firstYear3 <- crn[sss > 0.85][, .N, by = year][N >= 4, first(year)]
lastYear <- 2018
crnWide3 <- crn[year %in% firstYear3:lastYear,
                dcast(.SD, year ~ site, value.var = c('res'))]
X3 <- crnWide3[, -'year'] |> 
  as.matrix() |> 
  imputePCA(ncp = 6) |> 
  {\(x) x$completeObs}() 
PC3 <- prcomp(X3, scale. = TRUE)$x
sv <- c(1, 2)
sPC3 <- as.data.table(PC3[, sv, drop = FALSE])
lmFit3 <- PCR_reconstruction(PSepAug, sPC3, firstYear3, transform = 'none')
lmCV3 <- cvPCR(PSepAug, sPC3, firstYear3, transform = 'none',
              Z = cvFolds, use.robust.mean = FALSE)
recFinal3 <- lmFit3$rec
bcFit3 <- fitQmap(
  PSepAug$Qa,
  recFinal3[year %in% PSepAug$year, Q],
  'RQUANT',
  wet.day = FALSE)
recFinal3[, Qbc := doQmap(Q, bcFit)]
```

## Compare reconstructins from different SSS

```{r, fig.width=7, fig.height=6}
p1 <- ggplot() +
  geom_line(aes(year, Qbc, color = 'SSS = 0.6', linetype = 'SSS = 0.6'), 
            recFinal, alpha = 1) +
  geom_line(aes(year, Qbc, color = 'SSS = 0.75', linetype = 'SSS = 0.75'), 
            recFinal2, alpha = 1) +
  # scale_color_manual(values = MetBrewer::met.brewer('Kandinsky', 2))
  scale_color_manual(name = NULL, 
                     values = wesanderson::wes_palettes$Cavalcanti1[2:1]) +
  scale_linetype_manual(name = NULL, values = c(1, 2)) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    breaks = seq(200, 800, 100),
    minor_breaks = seq(200, 850, 50),
    limits = c(200, 850),
    guide = guide_prism_offset_minor()) +
  labs(x = NULL, y = 'Precipitation [mm]') +
  theme(
    legend.key.height = unit(1, 'cm'),
    legend.position = 'top')
p2 <- ggplot() +
  geom_line(aes(year, Qbc, color = 'SSS = 0.6', linetype = 'SSS = 0.6'), 
            recFinal, alpha = 1) +
  geom_line(aes(year, Qbc, color = 'SSS = 0.85', linetype = 'SSS = 0.85'), 
            recFinal3, alpha = 1) +
  scale_color_manual(name = NULL, 
                     values = wesanderson::wes_palettes$Cavalcanti1[c(2,5)]) +
  scale_linetype_manual(name = NULL, values = c(1, 2)) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1500, 2025, 25),
    limits = c(1500, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    breaks = seq(200, 800, 100),
    minor_breaks = seq(200, 850, 50),
    limits = c(200, 850),
    guide = guide_prism_offset_minor()) +
  labs(x = NULL, y = 'Precipitation [mm]') +
  theme(
    legend.key.height = unit(1, 'cm'),
    legend.position = 'top')
p1 / p2
```

## Compare skills

**Table S5**

```{r}
rbindlist(
  list(
    s1 = lmCV$metrics,
    s2 = lmCV2$metrics,
    s3 = lmCV3$metrics),
  idcol = 'cutoff')[, 1:4][, cutoff := c(0.6, 0.75, 0.85)][] |> 
  roundDT(3)
```

## Compare quantile regression results

```{r}
taus <- c(0.01, 0.05, 0.50, 0.95, 0.99)
names(taus) <- paste0(taus * 100, '%')
qrDT2 <- taus |> 
  lapply(\(tau) {
    fit <- rq(Qbc ~ year, tau, recFinal2)
    summ <- summary(fit, se = 'boot', R = 1000, bsmethod = 'wxy')
    data.table(
      b = fit$coefficients[1], 
      a = fit$coefficients[2], 
      signif = summ$coefficients[2, 4] < 0.1)
  }) |> 
  rbindlist(idcol = 'tau')

qrDT3 <- taus |> 
  lapply(\(tau) {
    fit <- rq(Qbc ~ year, tau, recFinal3)
    summ <- summary(fit, se = 'boot', R = 1000, bsmethod = 'wxy')
    data.table(
      b = fit$coefficients[1], 
      a = fit$coefficients[2], 
      signif = summ$coefficients[2, 4] < 0.1)
  }) |> 
  rbindlist(idcol = 'tau')
```

**Figure S4**

```{r, fig.width=7, fig.height=6}
p1 <- ggplot() +
  geom_line(aes(year, Qbc), recFinal2, color = 'gray85') +
  geom_abline(aes(slope = a, intercept = b, color = tau, linetype = signif), 
              qrDT2, show.legend = FALSE) +
  geom_line(aes(x = 1800, y = 400, color = tau, group = 1), qrDT2) +
  geom_point(aes(year, Qbc), recFinal2, color = 'gray50', shape = 1, size = 0.3) +
  scale_color_manual(
    values = c(reds[6], reds[3], 'darkviolet', blues9[4], blues9[7])) +
  scale_linetype_manual(values = c(2, 1)) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1625, 2025, 25),
    limits = c(1625, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    breaks = seq(200, 800, 100),
    minor_breaks = seq(200, 850, 50),
    limits = c(200, 850),
    guide = guide_prism_offset_minor()) +
  guides(color = guide_legend(
    title.position = 'top',
    title.hjust = 0.5,
    override.aes = list(size = 0.6))) +
  labs(x = NULL, y = 'Precipitation [mm]') +
  theme(
    legend.position = 'top')
p2 <- ggplot() +
  geom_line(aes(year, Qbc), recFinal3, color = 'gray85') +
  geom_abline(aes(slope = a, intercept = b, color = tau, linetype = signif), 
              qrDT3, show.legend = FALSE) +
  geom_line(aes(x = 1800, y = 400, color = tau, group = 1), qrDT3) +
  geom_point(aes(year, Qbc), recFinal3, color = 'gray50', shape = 1, size = 0.3) +
  scale_color_manual(
    values = c(reds[6], reds[3], 'darkviolet', blues9[4], blues9[7])) +
  scale_linetype_manual(values = c(2, 1)) +
  scale_x_continuous(
    expand = c(0, 0),
    minor_breaks = seq(1625, 2025, 25),
    limits = c(1625, 2025),
    guide = guide_prism_offset_minor()) +
  scale_y_continuous(
    breaks = seq(200, 800, 100),
    minor_breaks = seq(200, 850, 50),
    limits = c(200, 850),
    guide = guide_prism_offset_minor()) +
  guides(color = guide_legend(
    title.position = 'top',
    title.hjust = 0.5,
    override.aes = list(size = 0.6))) +
  labs(x = NULL, y = 'Precipitation [mm]') +
  theme(
    legend.position = 'top')
p1 / p2
```

## Using ARSTAN chronologies

```{r}
crnWide4 <- crn[year %in% firstYear:lastYear,
                dcast(.SD, year ~ site, value.var = c('ars'))]
X4 <- crnWide4[, -'year'] |> 
  as.matrix() |> 
  imputePCA(ncp = 6) |> 
  {\(x) x$completeObs}() 
PC4 <- prcomp(X4, scale. = TRUE)$x
sv <- c(1, 2)
sPC4 <- as.data.table(PC4[, sv, drop = FALSE])
lmFit4 <- PCR_reconstruction(PSepAug, sPC4, firstYear, transform = 'none')
lmCV4 <- cvPCR(PSepAug, sPC4, firstYear, transform = 'none',
              Z = cvFolds, use.robust.mean = FALSE)
recFinal4 <- lmFit4$rec
bcFit4 <- fitQmap(
  PSepAug$Qa,
  recFinal4[year %in% PSepAug$year, Q],
  'RQUANT',
  wet.day = FALSE)
recFinal4[, Qbc := doQmap(Q, bcFit)]
```

**Table S3**

```{r}
rbindlist(list(
  residual = lmCV$metrics,
  ARSTAN = lmCV4$metrics),
  idcol = 'Chronology')[, 1:4] |> 
  roundDT(3)
```
